### Set plot config

#### Specify plot styling variables

We store a separate configuration depending on output type of figure for easy switching between
output types. Only the currently selected output type variables in params + global variables for
plotting we may need later is returned. These variables are used to construct ggplot protoobjects
and can be accessed either with `plot_config()`, which returns all, or with
`plot_config("parameter")` to return a specific parameter.

```{r fun_plot_cfg}
plot_config <- function(variable = NULL, global_params = params) {
  p_cfg_vars <- list(
    global = list(
      axis_unit_color = "#4D4D4D" # used later when constructing axes from HTML
    ),
    poster = list(),
    paper = list(
      figure_dir = here(
        "output",
        "figures",
        if (isTRUE(global_params$telolength)) {
          paste0(global_params$hearing_var, "_telolength")
        } else {
          global_params$hearing_var
        }
      ),
      text_size_primary = 8,
      text_size_secondary = 9,
      signif_size = 14
    )
  )

  p_cfg_vars_comb <- c(p_cfg_vars$global, p_cfg_vars[[global_params$plot_output_type]])

  if (!is.null(variable)) p_cfg_vars_comb[[variable]] else p_cfg_vars_comb
}
```

#### Specify general plot theme

We first store the theme preset (theme_few), then theme overrides, then other theme items like
guides. All items specified in `plot_theme()` must be ggplot proto objects.

```{r fun_plot_theme}
plot_theme <- function(global_params = params) {
  config <- if (!is.null(global_params)) plot_config(global_params = global_params)

  list(
    theme_few(),
    theme(
      axis.text.x = element_markdown(
        "sans",
        size = config$text_size_primary
      ),
      axis.text.y = element_markdown(
        "sans",
        size = config$text_size_primary
      ),
      strip.text.x = element_markdown(
        "sans",
        size = config$text_size_secondary,
        lineheight = 1.1
      ),
      strip.text.y = element_markdown(
        "sans",
        size = config$text_size_secondary,
        lineheight = 1.1
      ),
      axis.title.x = element_markdown(
        "sans",
        size = config$text_size_secondary,
        lineheight = 1.1
      ),
      axis.title.y = element_markdown(
        "sans",
        size = config$text_size_secondary,
        lineheight = 1.1
      ),
      legend.title = element_blank(),
      legend.background = element_rect(fill = "transparent", color = NA),
      legend.box.background = element_rect(fill = "transparent", color = NA),
      legend.key = element_rect(fill = "transparent"),
      legend.position = "inside",
      legend.justification.inside = c(0, 1),
      legend.key.spacing.y = unit(1, "pt"),
      panel.border = element_part_rect(
        colour = "grey30",
        side = "lb",
        linewidth = 0.5
      ),
      plot.margin = margin(1, 1, 0, 1, unit = "pt")
    ),
    guides(
      colour = guide_legend(
        override.aes = list(size = 1.25),
        keywidth = unit(0.3, "cm"),
        keyheight = unit(0.3, "cm"),
        label.theme = element_markdown(
          "sans",
          size = config$text_size_secondary,
          margin = margin(l = 3, t = 1.5, unit = "pt")
        )
      )
    )
  )
}
```

#### Specify axis scales

```{r fun_plot_axes}
axes <- function(x = NULL, y = NULL, global_params = params) {
  config <- if (!is.null(global_params)) plot_config(global_params = global_params)

  if (!is.null(x)) {
    x <- switch(x,
      age = define_axis(
        config = config,
        axis = "x",
        type = "continuous",
        breaks = seq(20, 80, 10),
        limits = c(20, 82),
        name = "Age",
        unit = "Years"
      ),
      age_group = define_axis(
        config = config,
        axis = "x",
        type = "discrete",
        limits = c(1.4, 5.6),
        name = "Age",
        unit = "Years"
      ),
      age_accel_phenoage = define_axis(
        config = config,
        axis = "x",
        type = "continuous",
        breaks = seq(-20, 20, 20),
        limits = c(-20, 20),
        name = "PhenoAge Age Acceleration",
        unit = "PhenoAge – Age (Years)"
      ),
      age_accel_kdm = define_axis(
        config = config,
        axis = "x",
        type = "continuous",
        breaks = seq(-20, 20, 20),
        limits = c(-20, 20),
        name = "KDM Age Acceleration",
        unit = "KDM – Age (Years)"
      )
    )
  }

  if (!is.null(y)) {
    y <- switch(y,
      age_group = define_axis(
        config = config,
        axis = "y",
        type = "discrete",
        limits = c(1, 7),
        name = "Age",
        unit = "Years"
      ),
      hl = define_axis(
        config = config,
        axis = "y",
        type = "continuous",
        breaks = seq(0, 90, 30),
        limits = c(-10, 100),
        name = paste0(if (params$hearing_var == "FI_high") "High" else "Low", " Fletcher Index"),
        unit = "dB HL"
      ),
      hl_rate = define_axis(
        config = config,
        axis = "y",
        type = "continuous",
        breaks = seq(0, 2, 1),
        limits = c(-0.5, 2.6),
        name = "Rate of HL",
        unit = "dB HL/Year"
      ),
      hl_diff = define_axis(
        config = config,
        axis = "y",
        type = "continuous",
        breaks = seq(0, 2, 1),
        limits = c(-0.5, 2.6),
        name = "Deviation Rate of HL",
        unit = "dB HL/Year"
      ),
      dmse_bioage = define_axis(
        config = config,
        axis = "y",
        type = "continuous",
        breaks = seq(0, 8, 2),
        limits = c(-1, 8.5),
        name = "Additional Predictive Capacity",
        unit = "ΔMSE (dB HL)<sup>2</sup>"
      ),
      drmse_pred = define_axis(
        config = config,
        axis = "y",
        type = "continuous",
        breaks = seq(0, 0.25, 0.1),
        limits = c(-0.05, 0.25),
        name = "Additional Predictive Capacity",
        unit = "ΔRMSE (dB HL)"
      ),
      drmse_biomark = define_axis(
        config = config,
        axis = "y",
        type = "continuous",
        breaks = seq(0, 0.1, 0.05),
        limits = c(-0.025, 0.1),
        name = "Additional Predictive Capacity",
        unit = "ΔRMSE (dB HL)"
      ),
      dmse_conf = define_axis(
        config = config,
        axis = "y",
        type = "continuous",
        breaks = seq(0, 40, 5),
        limits = c(-1, 40),
        name = "Additional Predictive Capacity",
        unit = "ΔMSE (dB HL)<sup>2</sup>"
      ),
      add_pred_conf = define_axis(
        config = config,
        axis = "y",
        type = "continuous",
        breaks = seq(0, 14, 2),
        limits = c(-1, 15),
        name = "Additional Predictive Capacity",
        unit = "% Change MSE"
      )
    )
  }

  build_axes(x = x, y = y)
}
```

#### Specify colors

```{r fun_plot_colors}
color_scales <- function(color_list_name) {
  switch(color_list_name,
    bioage = build_color_scale(
      colors = c(
        "age" = "#000000",
        "age, hd_log" = "#256DAB",
        "age, kdm" = "#059748",
        "age, phenoage" = "#CB2027",
        "age, telolength" = if (isTRUE(params$telolength)) "#7B3A96" else NULL
      ),
      fill_colors = c(
        "age" = "#8C8C8C",
        "age, hd_log" = "#88BDE6",
        "age, kdm" = "#90CD97",
        "age, phenoage" = "#F07E6E",
        "age, telolength" = if (isTRUE(params$telolength)) "#B276B2" else NULL
      ),
      labels = c(
        "age" = "Age",
        "age, hd_log" = "Age + HD",
        "age, kdm" = "Age + KDM",
        "age, phenoage" = "Age + PhenoAge",
        "age, telolength" = if (isTRUE(params$telolength)) "Age + Telolength" else NULL
      )
    ),
    confounders = build_color_scale(
      colors = c(
        "full" = "#000000",
        "no_gender" = "#059748",
        "no_smoking" = "#9D722A",
        "no_race" = "#E5126F",
        "no_education" = "#DF5C24",
        "no_gender, race" = "#7B3A96",
        "full_with_phenoage" = "#256DAB"
      ),
      fill_colors = c(
        "full" = "#8C8C8C",
        "no_gender" = "#90CD97",
        "no_smoking" = "#BFA554",
        "no_race" = "#F6AAC9",
        "no_education" = "#FBB258",
        "no_gender, race" = "#BC99C7",
        "full_with_phenoage" = "#88BDE6"
      ),
      labels = c(
        "full" = "Full",
        "no_gender" = "Gender",
        "no_smoking" = "Smoking",
        "no_race" = "Race",
        "no_education" = "Education",
        "no_gender, race" = "Gender*Race",
        "full_with_phenoage" = "Phenoage"
      )
    ),
    phenoage = build_color_scale(
      colors = c(
        "age" = "#000000",
        "age, phenoage" = "#CB2027"
      ),
      fill_colors = c(
        "age" = "#8C8C8C",
        "age, phenoage" = "#F07E6E"
      ),
      labels = c(
        "age" = "Age",
        "age, phenoage" = "Age + PhenoAge"
      )
    ),
    gender = build_color_scale(
      colors = c(
        "1" = "#2986cc",
        "2" = "#c90076"
      ),
      fill_colors = c(
        "1" = "#6fa8dc",
        "2" = "#d5a6bd"
      ),
      labels = c(
        "1" = "Male",
        "2" = "Female"
      )
    ),
    race = build_color_scale(
      colors = c(
        "1" = "#E5126F",
        "3" = "#DF5C24",
        "4" = "#7B3A96",
        "5" = "#256DAB"
      ),
      fill_colors = c(
        "1" = "#F6AAC9",
        "3" = "#FBB258",
        "4" = "#BC99C7",
        "5" = "#88BDE6"
      ),
      labels = c(
        "1" = "Hispanic",
        "3" = "White",
        "4" = "Black",
        "5" = "Other"
      )
    ),
    biomarker_diff = build_color_scale(
      colors = c(
        "age, lncrp, phenoage" = "#CB2027",
        "age, lncrp" = "#059748",
        "age, phenoage" = "grey65"
      ),
      fill_colors = c(
        "age, lncrp, phenoage" = "#F07E6E",
        "age, lncrp" = "#90CD97",
        "age, phenoage" = "grey80"
      ),
      labels = c(
        "age, lncrp, phenoage" = "Age + CRP + PhenoAge",
        "age, lncrp" = "Age + CRP",
        "age, phenoage" = "Age + PhenoAge"
      ),
      breaks = c(
        "age, lncrp, phenoage",
        "age, lncrp",
        "age, phenoage"
      )
    ),
    best_biomarkers = build_color_scale(
      colors = c(
        "age, lncrp" = "#256DAB"
      ),
      fill_colors = c(
        "age, lncrp" = "#88BDE6"
      )
    )
  )
}
```

#### Specify geom style

Note that using `update_geom_defaults()` is global and requires `reset_geom_defaults()` to be reset.
`update_geom_defaults()` is only set once when switching to a different plot type and subsequent
plots with the same styling will not have repeated `update_geom_defaults()` calls to avoid overhead
and clutter. Plots must therefore be run in order.

```{r fun_plot_theme}
set_geom_theme <- function(geom_theme_name, global_params = params) {
  config <- if (!is.null(global_params)) plot_config(global_params = global_params)

  switch(geom_theme_name,
    ribbon = list(
      alpha = 0.15
    ),
    highlight_rect = list(
      alpha = 0.125,
      fill = "#ff0000"
    ),
    hline = list(
      linetype = "22",
      color = "grey20",
      linewidth = 0.2
    ),
    age_rate_point = list(
      alpha = 0.1,
      size = 0.5,
      shape = 1,
      stroke = 0.25,
      color = "grey30"
    ),
    age_rate_line = list(
      linewidth = 0.6,
      alpha = 1
    ),
    mse_point = list(
      size = 1.25,
      alpha = 1,
      shape = 16,
      stroke = 0.5
    ),
    mse_line = list(
      linewidth = 0.7,
      alpha = 1
    ),
    biomarker_linerange = list(
      linewidth = 0.75
    ),
    biomarker_point = list(
      size = 2.25,
      alpha = 1,
      shape = 16,
      stroke = 0.5
    )
  )
}
```

#### Set default geoms

The geom themes below are used in all plots. Plot-specific geoms are called on the first plot in a
sequence of plots that share the same geom theme.

```{r set_default_geoms}
update_geom_defaults("ribbon", set_geom_theme("ribbon"))
update_geom_defaults("rect", set_geom_theme("highlight_rect"))
update_geom_defaults("hline", set_geom_theme("hline"))
```

### Rate of aging panel

#### Main panel (paper)

```{r plot_aging, warning=FALSE}
update_geom_defaults("point", set_geom_theme("age_rate_point"))
update_geom_defaults("line", set_geom_theme("age_rate_line"))

p0 <- ggplot() +
  geom_point(
    data = df,
    aes(
      x = age,
      y = .data[[params$hearing_var]]
    )
  ) +
  geom_line(
    data = output_age_rate$predictions$linear,
    aes(
      x = age,
      y = estimate
    )
  ) +
  geom_ribbon(
    data = output_age_rate$predictions$linear,
    aes(
      x = age,
      ymin = ci_lower,
      ymax = ci_upper
    )
  ) +
  axes(
    x = "age",
    y = "hl"
  ) +
  plot_theme() +
  theme(
    plot.margin = margin(1, 0, 0, 1, unit = "pt"),
    axis.title.x = element_blank()
  )

p1 <- ggplot() +
  geom_point(
    data = df,
    aes(
      x = age,
      y = .data[[params$hearing_var]]
    )
  ) +
  geom_line(
    data = output_age_rate$predictions$quadratic,
    aes(
      x = age,
      y = estimate
    )
  ) +
  geom_ribbon(
    data = output_age_rate$predictions$quadratic,
    aes(
      x = age,
      ymin = ci_lower,
      ymax = ci_upper
    )
  ) +
  axes(
    x = "age",
    y = "hl"
  ) +
  plot_theme() +
  theme(
    plot.margin = margin(1, 0, 0, 1, unit = "pt"),
    axis.title.y = element_blank(),
    axis.title.x = element_blank()
  )

p2 <- ggplot() +
  geom_rect(
    data = pwc_age_rate_boundaries,
    aes(
      xmin = x_min,
      xmax = x_max,
      ymin = -Inf,
      ymax = Inf
    )
  ) +
  geom_point(
    data = df,
    aes(
      x = age,
      y = .data[[params$hearing_var]]
    )
  ) +
  geom_ribbon(
    data = output_age_rate$predictions$smooth,
    aes(
      x = age,
      ymin = ci_lower + coef(output_age_rate$models$smooth)["(Intercept)"],
      ymax = ci_upper + coef(output_age_rate$models$smooth)["(Intercept)"]
    )
  ) +
  geom_line(
    data = output_age_rate$predictions$smooth,
    aes(
      x = age,
      y = estimate + coef(output_age_rate$models$smooth)["(Intercept)"]
    )
  ) +
  axes(
    x = "age",
    y = "hl"
  ) +
  plot_theme() +
  theme(
    plot.margin = margin(1, 1, 0, 0, unit = "pt"),
    axis.title.y = element_blank(),
    axis.title.x = element_blank()
  )

p3 <- ggplot() +
  geom_hline(yintercept = 0) +
  geom_line(
    data = output_age_rate$predictions$linear_derivative,
    aes(
      x = age,
      y = derivative
    )
  ) +
  geom_ribbon(
    data = output_age_rate$predictions$linear_derivative,
    aes(
      x = age,
      ymin = ci_lower,
      ymax = ci_upper
    )
  ) +
  axes(
    x = "age",
    y = "hl_rate"
  ) +
  plot_theme() +
  theme(plot.margin = margin(0, 0, 0, 1, unit = "pt"))

p4 <- ggplot() +
  geom_hline(yintercept = 0) +
  geom_line(
    data = output_age_rate$predictions$quadratic_derivative,
    aes(
      x = age,
      y = derivative
    )
  ) +
  geom_ribbon(
    data = output_age_rate$predictions$quadratic_derivative,
    aes(
      x = age,
      ymin = ci_lower,
      ymax = ci_upper
    )
  ) +
  axes(
    x = "age",
    y = "hl_rate"
  ) +
  plot_theme() +
  theme(
    plot.margin = margin(0, 0, 0, 1, unit = "pt"),
    axis.title.y = element_blank()
  )

p5 <- ggplot() +
  geom_rect(
    data = pwc_age_rate_boundaries,
    aes(
      xmin = x_min,
      xmax = x_max,
      ymin = -Inf,
      ymax = Inf
    )
  ) +
  geom_hline(yintercept = 0) +
  geom_ribbon(
    data = output_age_rate$predictions$derivative_smooth,
    aes(
      x = age,
      ymin = ci_lower,
      ymax = ci_upper
    )
  ) +
  geom_line(
    data = output_age_rate$predictions$derivative_smooth,
    aes(
      x = age,
      y = derivative
    )
  ) +
  axes(
    x = "age",
    y = "hl_rate"
  ) +
  plot_theme() +
  theme(
    plot.margin = margin(0, 1, 0, 0, unit = "pt"),
    axis.title.y = element_blank()
  )

p6 <- ggplot() +
  geom_rect(
    data = pwc_age_rate_boundaries,
    aes(
      xmin = x_min,
      xmax = x_max,
      ymin = -Inf,
      ymax = Inf
    )
  ) +
  geom_hline(yintercept = 0) +
  geom_ribbon(
    data = output_age_rate$predictions$derivative_smooth,
    aes(
      x = age,
      ymin = ci_lower_adj,
      ymax = ci_upper_adj
    )
  ) +
  geom_line(
    data = output_age_rate$predictions$derivative_smooth,
    aes(
      x = age,
      y = derivative_adj
    )
  ) +
  axes(
    x = "age",
    y = "hl_diff"
  ) +
  plot_theme() +
  theme(plot.margin = margin(0, 0, 1, 0, unit = "pt"))


p <- align_plots(p0, p1, p2, p3, p4, p5, p6, align = "hv")

p <- plot_grid(
    plot_grid(
      p[[1]], NULL,
      p[[2]], NULL,
      p[[3]],
      labels = c("A", "", "B", "", "C"),
      nrow = 1,
      rel_widths = c(1, -0.01, 1, -0.01, 1),
      label_x = 0.04,
      label_y = 1.02
    ),
    NULL,
    plot_grid(
      p[[4]], NULL,
      p[[5]], NULL,
      p[[6]],
      labels = c("D", "", "E", "", "F"),
      nrow = 1,
      rel_widths = c(1, -0.01, 1, -0.01, 1),
      label_x = 0.04,
      label_y = 1.02
    ),
  NULL,
  plot_grid(
    NULL, p[[7]], NULL,
    nrow = 1,
    rel_widths = c(0.25, 0.5, 0.25),
    labels = c("", "G", ""),
    label_x = 0.04,
    label_y = 1.02
  ),
  ncol = 1,
  rel_heights = c(1, 0, 1, 0, 1)
)

sameplot(
  p,
  file = here(plot_config("figure_dir"), "age_rate_panel.png"),
  width = params$double_column,
  height = 6,
  res = 600,
  save = params$save_figures
)
```
